
My deepest gratitude is to my advisor Professor Yanming Di. I have been extremely fortunate to have 
an advisor who guided and supported me in my PhD study/research with great patience. Yanming taught 
me to think ``science", think ``positive", and think ``why", which in my opinion is the essential 
attitude towards scientific research. He also taught me how to write research articles clearly and 
logically. These are the most important skills that I've learned during my PhD training and will 
benefit me greatly in my future career. He helped me identify interesting research questions and 
guided me carefully when my steps faltered. His encouragement and help have led me all the way 
towards finishing this dissertation.

I'm very grateful to my committee members: Dr. Duo Jiang (Statistics), Dr. Sarah Emerson 
(Statistics),  Dr. Jeff Chang (Botany and Plant Pathology) and Dr. Yuan Jiang (Statistics). Duo 
helped me throughout the research work of Chapter \ref{chap3}, from developing method to writing 
the manuscript. I'm indebted to her for her generous advice and encouragement. Sarah provided 
insightful comments and constructive criticisms at different stages of my research, which helped me 
understand the technical details and enrich my ideas. I thank Jeff for raising biological research 
questions and providing biological background support/explanation, and Yuan for helpful discussions 
and suggestions. 

I would like to thank all the faculty and staff in the Department of Statistics, for teaching 
the courses and providing financial support during my pursuit of a doctorate degree. Special 
thanks to Professor Lisa Madsen for offering valuable advice and help during my time at Oregon 
State University. 

My thanks also go out to the Cosine IT Services at College of Science for providing high 
performance computing resources, and to Dr. Shawn O'Neil and Casey Dinsmore for their technical 
support on Linux cluster computing. 

Special thanks are extended to those who helped me out during this challenging period at Corvallis. 
My deep appreciation goes out to my roommate Dr. Weimin Liu, who is now an assistant professor at 
ShanghaiTech University. For the past thirty months, he has been a mentor for my everyday life. His 
enthusiasm for academia and devotion to his research work motivated me to persist when I felt 
frustrated. He cared about my thinking, heard my voice, and helped me see the pros and cons when I 
was in a dilemma. Thanks go out to Yuanli Hu, Qiang Dai, Luna Sun, Gu Mi, Hai Yang, Lu Wang, Yandan 
Yang, the groups of friends ``Buddies at OSU" and ``Muscle-Meat-Heads". I am indebted to all my 
other friends (not listed here) who have been and will always be very helpful and supportive in my 
life.


Finally, I must express my profound gratitude to my family. I thank my parents for giving me the 
freedom to choose my own way of life, and for believing in and being supportive of the choices I 
made at every stage.  I thank my brother for providing me unfailing support and continuous 
encouragement throughout my years of study. %I welcome my nephew (or niece), the newly born baby, 
%to join my family.



